---
title: "CUNY SPS DATA 606 Homework 2"
author: "Betsy Rosalen"
date: "February 18, 2018"
output:
  html_document:
    theme: cerulean
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(openintro)
data(iris)
data(heartTr)
```

## Exercise 2.6

If you roll a pair of fair dice, what is the probability of

#####  (a) getting a sum of 1?

The probability of getting a sum of one when rolling a pair of fair dice is 0 since the minimum sum would be 2 if you rolled two 1's.

#####  (b) getting a sum of 5?

The probability of getting a sum of 5 when rolling a pair of fair dice is $\frac{4}{36}$ since there are 4 ways you can get a sum of 5 (1, 4), (2, 3), (3, 2), and (4, 1) out of 36 possible rolls.

```{r}
# There are 6 times 6 combinations possible when rolling two 6-sided dice
6*6
```


#####  (c) getting a sum of 12?

The probability of getting a sum of 12 when rolling a pair of fair dice is $\frac{1}{36}$ since there is only one way to get a sum of 12 if you rolled two 6's.

## Exercise 2.8

The American Community Survey is an ongoing survey that provides data every year to give communities the current information they need to plan investments and services. The 2010 American Community Survey estimates that 14.6% of Americans live below the poverty line, 20.7% speak a language other than English (foreign language) at home, and 4.2% fall into both categories.

#####  (a) Are living below the poverty line and speaking a foreign language at home disjoint?

No, since there are 4.2% of the population that fall into both categories, living below the poverty line, and speaking a language other than English at home are not disjoint.

#####  (b) Draw a Venn diagram summarizing the variables and their associated probabilities.



#####  (c) What percent of Americans live below the poverty line and only speak English at home?

10.4% of Americans (14.6% minus 4.2%) live below the poverty line and only speak English at home.

```{r}
14.6 - 4.2
```

#####  (d) What percent of Americans live below the poverty line or speak a foreign language at home?

31.1% of Americans live below the poverty line or speak a foreign language at home.

```{r}
14.6 + 20.7 - 4.2
```

#####  (e) What percent of Americans live above the poverty line and only speak English at home?

68.9% of Americans live above the poverty line and only speak English at home. 

```{r}
100 - 14.6 - (20.7 - 4.2)
```

#####  (f) Is the event that someone lives below the poverty line independent of the event that the person speaks a foreign language at home? 

If the two events were independent then knowing the outcome of one should not affect the probability of the second, but the probability of living below the poverty line is significantly higher when you know that they speak a language other than English at home.  As we saw above 10.4 percent of Americans live below the poverty line and only speak English at home and we know that 4.2 percent live below the poverty line and speak a foreign language at home. Knowing that you speak a foreign language at home means there is a 20.2% probability that live below the poverty line wherease if you do not speak a foreign language at home, your probability of living below the poverty line is only 13.1%.

```{r}
4.2/20.7
10.4/(100-20.7)
```


## Exercise 2.20

Assortative mating is a nonrandom mating pattern where individuals with similar genotypes and/or phenotypes mate with one another more frequently than what would be expected under a random mating pattern. Researchers studying this topic collected data on eye colors of 204 Scandinavian men and their female partners. The table below summarizes the results. For simplicity, we only include heterosexual relationships in this exercise.

Partner (female) runs across horizontally and Self (male) runs vertically

--- | Blue | Brown | Green | Total
----- | ----- | ----- | ----- | -----
Blue | 78 | 23 | 13 | 114
Brown | 19 | 23 | 12 | 54
Green | 11 | 9 | 16 | 36
Total | 108 | 55 | 41 | 204

#####  (a) What is the probability that a randomly chosen male respondent or his partner has blue eyes?

```{r}
# probability that a randomly chosen male respondent or his partner has blue eyes
(108 + 114 - 78)/204
```

#####  (b) What is the probability that a randomly chosen male respondent with blue eyes has a partner with blue eyes?

```{r}
# probability that a randomly chosen male respondent with blue eyes has a partner with blue eyes
78/114
```

#####  (c.1) What is the probability that a randomly chosen male respondent with brown eyes has a partner with blue eyes? 

```{r}
# probability that a randomly chosen male respondent with brown eyes has a partner with blue eyes
19/54
```

##### (c.2) What about the probability of a randomly chosen male respondent with green eyes having a partner with blue eyes?

```{r}
# probability of a randomly chosen male respondent with green eyes has a partner with blue eyes
11/36
```

#####  (d) Does it appear that the eye colors of male respondents and their partners are independent?  Explain your reasoning.

No, it appears that male respondents eye color and their partner's eye color are not independent.  The probability of a partner's eye color being the same as their own is higher than the probability of any other color in all cases.

```{r}
# probability that a randomly chosen male respondent with brown eyes has a partner with brown eyes
23/54
# probability of a randomly chosen male respondent with green eyes has a partner with green eyes
16/36
```


## Exercise 2.30

The table below shows the distribution of books on a bookcase based on whether they are nonfiction or fiction and hardcover or paperback.

--- | Hardcover | Paperback | Total
----- | ----- | ----- | -----
Fiction | 13 | 59 | 72
Nonfiction | 15 | 8 | 23
Total | 28 | 67 | 95

#####  (a) Find the probability of drawing a hardcover book first then a paperback fiction book second when drawing without replacement.

```{r}
28/95 * 59/94
```

#####  (b) Determine the probability of drawing a fiction book first and then a hardcover book second, when drawing without replacement.

```{r}
# if the first book was a hardcover fiction book
72/95 * 27/94

# if the first book was a paperback fiction book
72/95 * 28/94
```

#####  (c) Calculate the probability of the scenario in part (b), except this time complete the calculations under the scenario where the first book is placed back on the bookcase before randomly drawing the second book.

```{r}
# if the first book was a hardcover fiction book
72/95 * 28/95
```

#####  (d) The final answers to parts (b) and (c) are very similar. Explain why this is the case. 

From the textbook page 103: "When the sample size is only a small fraction of the population (under 10%), observations are nearly independent even when sampling without replacement."  In this case our sample is only 2/95 or about 2% of our population.

```{r}
2/95
```


## Exercise 2.38

An airline charges the following baggage fees: \$25 for the first bag and \$35 for the second. Suppose 54% of passengers have no checked luggage, 34% have one piece of checked luggage and 12% have two pieces. We suppose a negligible portion of people check more than two bags.

#####  (a) Build a probability model, compute the average revenue per passenger, and compute the corresponding standard deviation.

```{r}

```


#####  (b) About how much revenue should the airline expect for a flight of 120 passengers? With what standard deviation? Note any assumptions you make and if you think they are justified. 

```{r}

```


## Exercise 2.44

The relative frequency table below displays the distribution of annual total personal income (in 2009 inflation-adjusted dollars) for a representative sample of 96,420,486 Americans. These data come from the American Community Survey for 2005-2009. This sample is comprised of 59% males and 41% females.

Income | Total
----- | -----
$1 to $9,999 or loss | 2.2%
$10,000 to $14,999 | 4.7%
$15,000 to $24,999 | 15.8%
$25,000 to $34,999 | 18.3%
$35,000 to $49,999 | 21.2%
$50,000 to $64,999 | 13.9%
$65,000 to $74,999 | 5.8%
$75,000 to $99,999 | 8.4%
$100,000 or more | 9.7%

#####  (a) Describe the distribution of total personal income.

The distribution seems to be slightly right skewed with the mode below the halfway point.

#####  (b) What is the probability that a randomly chosen US resident makes less than $50,000 per year?

```{r}
lessthan50k <- (21.2 + 18.3 + 15.8 + 4.7 + 2.2)
lessthan50k
```

#####  (c) What is the probability that a randomly chosen US resident makes less than $50,000 per year and is female? Note any assumptions you make.

Assuming that the proportion of males and females in each income bracket are the same as the proportion of males to females in the overall population the probability that a randomly chosen US resident makes less than $50,000 per year and is female is 25.5%.

```{r}
lessthan50k * .41
```

#####  (d) The same data source indicates that 71.8% of females make less than $50,000 per year. Use this value to determine whether or not the assumption you made in part (c) is valid.

Since the overall proportion of the population who make less than \$50,000 per year is 62.2% we would expect that if the proprtions of males and females at each income bracket were the same as the population as a whole then 62.2 percent of women should make below \$50,000 per year.  So my assumption above is not valid.


