---
title: "DATA 606 Notes - Week 1"
output: html_document
---

# Jan 31, 2018

## The DATA606 R Package

The package can be installed from Github using the devtools package.

`devtools::install_github('jbryer/DATA606')`

### Important Functions

Function | Description
----- | -----
`devtools::install_github('jbryer/DATA606')` | Install package from Github using devtools
`library('DATA606')` | Load the package
`vignette(package='DATA606')` | Lists vignettes in the DATA606 package
`vignette('os3')` | Loads a PDF of the OpenIntro Statistics book
`data(package='DATA606')` | Lists data available in the package
`getLabs()` | Returns a list of the available labs
`viewLab('Lab0')` | Opens Lab0 in the default web browser
`startLab('Lab0')` | Starts Lab0 (copies to getwd()), opens the Rmd file
`shiny_demo()` | Lists available Shiny apps
`getwd()` | get the working directory
`setwd()` | get the working directory

R Markdown files are provided for all the labs. You can start a lab using the `DATA606::startLab` function.

However, creating new R Markdown files in RStudio can be done by clicking File > New File > R Markdown.

In RStudio, you can also set the working directory on the Files tab by clicking More, then Set as Working Directory.

## Intro to Data

We will use the lego R package in this class which contains information about every Lego set manufactured from 1970 to 2014, a total of 5710 sets.

```{r}
#devtools::install_github("seankross/lego")
library(lego)
data(legosets)
```

## Contingency Tables

```{r}
table(legosets$Availability, useNA='ifany')
```

Creates a table of totals for the availabiliy variable in the legosets dataframe - Part of the Base R package

```{r}
table(legosets$Availability, legosets$Packaging, useNA='ifany')
```

Creates a table of totals for the availabiliy variable in the legosets dataframe for each packaging option

`useNA='ifany'` - Will ensure any missing values are counted

## Proportional Tables

```{r}
prop.table(table(legosets$Availability))
``` 

Creates a table of proportions for the availabiliy variable in the legosets dataframe - Part of the Base R package

## Bar Plots

```{r}
barplot(table(legosets$Availability), las=3)
```
```{r}
barplot(prop.table(table(legosets$Availability)), las=3)
```

## Mosaic Plot

```{r}
library(vcd)
mosaic(HairEyeColor, shade=TRUE, legend=TRUE)
```

## Measures of Center

```{R}
mean(legosets$Pieces, na.rm=TRUE)
median(legosets$Pieces, na.rm=TRUE)
```

## Measures of Spread

```{R}
var(legosets$Pieces, na.rm=TRUE)
sqrt(var(legosets$Pieces, na.rm=TRUE))
sd(legosets$Pieces, na.rm=TRUE)
fivenum(legosets$Pieces, na.rm=TRUE)
IQR(legosets$Pieces, na.rm=TRUE)
```

## The summary Function

```{R}
summary(legosets$Pieces)
```

## The psych Package

```{R}
library(psych)
describe(legosets$Pieces, skew=FALSE)
```

```{R}
describeBy(legosets$Pieces, group = legosets$Availability, skew=FALSE, mat=TRUE)
```

## Dot Plot

```{R}
stripchart(legosets$Pieces)
```

```{R}
par.orig <- par(mar=c(1,10,1,1))
stripchart(legosets$Pieces ~ legosets$Availability, las=1)
```

## Histograms

```{R}
hist(legosets$Pieces)
```

## Transformations
With highly skewed distributions, it is often helpful to transform the data. The log transformation is a common approach, especially when dealing with salary or similar data.

```{R}
hist(log(legosets$Pieces))
```

## Density Plots

```{R}
plot(density(legosets$Pieces, na.rm=TRUE), main='Lego Pieces per Set')
```

## Box Plots

```{R}
boxplot(legosets$Pieces)
```

## Scatter Plots

```{R}
plot(legosets$Pieces, legosets$USD_MSRP)
```

## Examining Possible Outliers 
### expensive sets

```{R}
legosets[which(legosets$USD_MSRP >= 400),]
```

### big sets

```{R}
legosets[which(legosets$Pieces >= 4000),]
```

```{R}
plot(legosets$Pieces, legosets$USD_MSRP)
bigAndExpensive <- legosets[which(legosets$Pieces >= 4000 | legosets$USD_MSRP >= 400),]
text(bigAndExpensive$Pieces, bigAndExpensive$USD_MSRP, labels=bigAndExpensive$Name)
```

## Robust Statistics

Median and IQR are more robust to skewness and outliers than mean and SD. Therefore,

- for skewed distributions it is often more helpful to use median and IQR to describe the center and spread
- for symmetric distributions it is often more helpful to use the mean and SD to describe the center and spread

## Just say NO to pie charts!

"There is no data that can be displayed in a pie chart that cannot better be displayed in some other type of chart"
 --John Tukey