---
title: 'Inference for numerical data'
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

## North Carolina births

In 2004, the state of North Carolina released a large data set containing information on births recorded in this state. This data set is useful to researchers studying the relation between habits and practices of expectant mothers and the birth of their children. We will work with a random sample of observations from this data set.

## Exploratory analysis

Load the `nc` data set into our workspace.

```{r load-data, eval=TRUE}
load("more/nc.RData")
```

We have observations on 13 different variables, some categorical and some numerical. The meaning of each variable is as follows.

variable         | description
---------------- | -----------
`fage`           | father's age in years.
`mage`           | mother's age in years.
`mature`         | maturity status of mother.
`weeks`          | length of pregnancy in weeks.
`premie`         | whether the birth was classified as premature (premie) or full-term.
`visits`         | number of hospital visits during pregnancy.
`marital`        | whether mother is `married` or `not married` at birth.
`gained`         | weight gained by mother during pregnancy in pounds.
`weight`         | weight of the baby at birth in pounds.
`lowbirthweight` | whether baby was classified as low birthweight (`low`) or not (`not low`).
`gender`         | gender of the baby, `female` or `male`.
`habit`          | status of the mother as a `nonsmoker` or a `smoker`.
`whitemom`       | whether mom is `white` or `not white`.

1.  What are the cases in this data set? How many cases are there in our sample?
    ```{r}
    dim(nc)
    ```
    
    Each birth is a case and there are 1000 cases in the sample.  

As a first step in the analysis, we should consider summaries of the data. This
can be done using the `summary` command:

```{r summary, eval=TRUE}
summary(nc)
```

As you review the variable summaries, consider which variables are categorical and which are numerical. For numerical variables, are there outliers? If you aren't sure or want to take a closer look at the data, make a graph.

Consider the possible relationship between a mother's smoking habit and the weight of her baby. Plotting the data is a useful first step because it helps us quickly visualize trends, identify strong associations, and develop research questions.

2.  Make a side-by-side boxplot of `habit` and `weight`. What does the plot highlight about the relationship between these two variables?
    ```{r}
    boxplot(nc$weight ~ nc$habit)
    ```
    
    The birth weight for children born to non-smokers seems to be slightly higher on average with a higher median and a larger total range, but a slightly smaller interquartile range.  Also, there are more outliers on both ends of the range for non-smokers than for smokers.

The box plots show how the medians of the two distributions compare, but we can also compare the means of the distributions using the following function to split the `weight` variable into the `habit` groups, then take the mean of each using the `mean` function.

```{r by-means, eval=TRUE}
by(nc$weight, nc$habit, mean)
```

There is an observed difference, but is this difference statistically significant? In order to answer this question we will conduct a hypothesis test
.

## Inference

3.  Check if the conditions necessary for inference are satisfied. Note that you will need to obtain sample sizes to check the conditions. You can compute the group size using the same `by` command above but replacing `mean` with `length`.

    ```{r by-length, eval=TRUE}
    by(nc$weight, nc$habit, length)
    ```
    ```{r}
    par(mfrow=c(1,2))
    hist(nc$weight[nc$habit == "nonsmoker"], main = "Non-smokers", xlab = "Birth Weight")
    hist(nc$weight[nc$habit == "smoker"], main = "Smokers", xlab = "Birth Weight")
    ```
    

    Assuming that our total sample was randomly selected and is less than 10% of the population, both the smoker and non-smoker subsets of our sample can be assumed to be independent.  Although there is a slight left skew in the distribution for both groups since they each have more than 30 cases the slight skew can safely be ignored and the normality condition is met.  Both groups meet the conditions for inference.  
    
4.  Write the hypotheses for testing if the average weights of babies born to smoking and non-smoking mothers are different.

    $H_0$: The average weights of babies born to smoking and non-smoking mothers are the same.
    
    $$\mu_{smoker} = \mu_{non-smoker}$$

    $H_A$: The average average weights of babies born to smoking and non-smoking mothers are different.
    
    $$\mu_{smoker} \neq \mu_{non-smoker}$$

Next, we introduce a new function, `inference`, that we will use for conducting hypothesis tests and constructing confidence intervals. 

```{r inf-weight-habit-ht, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```

Let's pause for a moment to go through the arguments of this custom function. The first argument is `y`, which is the response variable that we are interested in: `nc$weight`. The second argument is the explanatory variable, `x`, which is the variable that splits the data into two groups, smokers and non-smokers: `nc$habit`. The third argument, `est`, is the parameter we're interested in: `"mean"` (other options are `"median"`, or `"proportion"`.) Next we decide on the `type` of inference we want: a hypothesis test (`"ht"`) or a 
confidence interval (`"ci"`). When performing a hypothesis test, we also need to supply the `null` value, which in this case is `0`, since the null hypothesis sets the two population means equal to each other. The `alternative` hypothesis can be `"less"`, `"greater"`, or `"twosided"`. Lastly, the `method` of inference can be `"theoretical"` or `"simulation"` based.

5.  Change the `type` argument to `"ci"` to construct and record a confidence interval for the difference between the weights of babies born to smoking and non-smoking mothers.

    ```{r eval=TRUE, tidy=FALSE}
    inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", null = 0, 
              alternative = "twosided", method = "theoretical")
    ```

By default the function reports an interval for ($\mu_{nonsmoker} - \mu_{smoker}$). We can easily change this order by using the `order` argument:

```{r inf-weight-habit-ci, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", 
          order = c("smoker","nonsmoker"))
```

* * *

## On your own

-   Calculate a 95% confidence interval for the average length of pregnancies 
(`weeks`) and interpret it in context. Note that since you're doing inference 
on a single population parameter, there is no explanatory variable, so you can 
omit the `x` variable from the function.

    ```{r eval=TRUE, tidy=FALSE}
    inference(y = nc$weeks, est = "mean", type = "ci", null = 0, 
              alternative = "twosided", method = "theoretical")
    ```
    
    We can be 95% confident that the average length of pregnancies for the population is between 38.1528 and 38.5165 weeks.
    
-   Calculate a new confidence interval for the same parameter at the 90% 
confidence level. You can change the confidence level by adding a new argument 
to the function: `conflevel = 0.90`.

    ```{r eval=TRUE, tidy=FALSE}
    inference(y = nc$weeks, est = "mean", type = "ci", null = 0, 
              alternative = "twosided", method = "theoretical", conflevel = 0.90)
    ```

    We can be 90% confident that the average length of pregnancies for the population is between 38.182 and 38.4873 weeks.

-   Conduct a hypothesis test evaluating whether the average weight gained by 
younger mothers is different than the average weight gained by mature mothers.

    ```{r}
    inference(y = nc$gained, x=nc$mature, est = "mean", type = "ht", null = 0, 
              alternative = "twosided", method = "theoretical")
    ```

    Since our p-value is 0.1686 the probability of observing this difference if the null hypothesis is true is 0.1686 which exceeds our signifgance level of 0.05, so we fail to reject the null hypothesis. There is not enough statistical evidence to suggest that there is a difference in weight gain between mature mothers and younger mothers.

-   Now, a non-inference task: Determine the age cutoff for younger and mature 
mothers. Use a method of your choice, and explain how your method works.

    ```{r}
    by(nc$mage, nc$mature, range)
    ```

    I used the `by` function introduced above to find the range in age of both mature and younger moms.  Since the max age for younger moms is 34 and the minimum age for mature mothers is 35, the cutoff between the two groups is between 34 and 35 years of age.

-   Pick a pair of numerical and categorical variables and come up with a 
research question evaluating the relationship between these variables. 
Formulate the question in a way that it can be answered using a hypothesis test
and/or a confidence interval. Answer your question using the `inference` 
function, report the statistical results, and also provide an explanation in 
plain language.

    #### Research question: 
    
    Is the amount of weight gained by the mother during pregnancy different for births classified as premature vs. births classified as full term?
    
    Variables needed for analysis:

    variable         | description
    ---------------- | -----------
    `premie`         | whether the birth was classified as premature (premie) or full-term.
    `gained`         | weight gained by mother during pregnancy in pounds.

    #### Hypotheses:

    $H_0$: The average weight gained by the mother during pregnancy is the same for both full-term and premie births.
    
    $$\mu_{full-term} = \mu_{premie}$$

    $H_A$: The average weight gained by the mother during pregnancy is not the same for full-term and premie births.
    
    $$\mu_{full-term} \neq \mu_{premie}$$

    ```{r eval=TRUE, tidy=FALSE}
    inference(y = nc$gained, x = nc$premie, est = "mean", type = "ht", null = 0, 
              alternative = "twosided", method = "theoretical")
    ```
    
    If the null hypothesis is true, the probability of observing this difference in the mean weight gained by mothers of births classified as full-term vs. those classified as premature is 0.  We reject the null hypothesis in favor of the alternative huypothesis. There is significant statistical evidence to suggest that there is a difference in weight gain between mothers of births classified as full-term vs. those classified as premature.

    
   
<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was adapted for OpenIntro by Mine &Ccedil;etinkaya-Rundel from a lab 
written by the faculty and TAs of UCLA Statistics.
</div>
