---
title: 'Inference for categorical data'
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

In August of 2012, news outlets ranging from the [Washington
Post](http://www.washingtonpost.com/national/on-faith/poll-shows-atheism-on-the-rise-in-the-us/2012/08/13/90020fd6-e57d-11e1-9739-eef99c5fb285_story.html) to the [Huffington
Post](http://www.huffingtonpost.com/2012/08/14/atheism-rise-religiosity-decline-in-america_n_1777031.html)
ran a story about the rise of atheism in America. The source for the story was 
a poll that asked people, "Irrespective of whether you attend a place of 
worship or not, would you say you are a religious person, not a religious 
person or a convinced atheist?" This type of question, which asks people to 
classify themselves in one way or another, is common in polling and generates 
categorical data. In this lab we take a look at the atheism survey and explore 
what's at play when making inference about population proportions using 
categorical data.

## The survey

To access the press release for the poll, conducted by WIN-Gallup 
International, click on the following link:

*<https://github.com/jbryer/EPSY530Summer2015/blob/master/Labs/Global_INDEX_of_Religiosity_and_Atheism_PR__6.pdf>*

Take a moment to review the report then address the following questions.

1.  In the first paragraph, several key findings are reported. Do these 
    percentages appear to be *sample statistics* (derived from the data 
    sample) or *population parameters*?
    
    The statistics reported in the first paragraph are *sample statistics* (derived from the data sample) since it says, "according to the latest global poll".

2.  The title of the report is "Global Index of Religiosity and Atheism". To
    generalize the report's findings to the global human population, what must 
    we assume about the sampling method? Does that seem like a reasonable 
    assumption?
    
    To generalize the results of the poll to the global human population we must assume that the survey respondents were chosen at random from the population and that there was no bias in the sample.  It seems unlikely to me that there is no non-response bias, or convenience bias in a poll that is global in scale.  Since there are 195 countries in the world and the survey only sampled from 57 countries, and there are 6 continents (if you exclude Antartica) and the survey only sampled 5, I don't think you could generalize the results to the global population as a whole, only possibly to each individual country that was sampled. Even that is a stretch since the summary at the end says that 11 out of the 57 countries were only polled in urban areas.

## The data

Turn your attention to Table 6 (pages 15 and 16), which reports the
sample size and response percentages for all 57 countries. While this is
a useful format to summarize the data, we will base our analysis on the
original data set of individual responses to the survey. Load this data
set into R with the following command.

```{r head-data, eval=TRUE}
load("more/atheism.RData")
```

3.  What does each row of Table 6 correspond to? What does each row of 
    `atheism` correspond to?
    
    Each row in table 6 corresponds to the aggragated results from one country that was polled.  
    
    Each row in the `aetheism` dataset corresponds to one respondent in the poll, and includes the country they are from, whether they are an aetheist or not and the year they were polled.

To investigate the link between these two ways of organizing this data, take a 
look at the estimated proportion of atheists in the United States. Towards 
the bottom of Table 6, we see that this is 5%. We should be able to come to 
the same number using the `atheism` data.

4.  Using the command below, create a new dataframe called `us12` that contains
    only the rows in `atheism` associated with respondents to the 2012 survey 
    from the United States. Next, calculate the proportion of atheist 
    responses. Does it agree with the percentage in Table 6? If not, why?

    ```{r us-atheism, eval=TRUE}
us12 <- subset(atheism, nationality == "United States" & year == "2012")
us12$nationality <- factor(us12$nationality)
prop.table(table(us12$response))
```

    Yes, it does agree with the percentage in tabe 6 if you round to two digits.

## Inference on proportions

As was hinted at in Exercise 1, Table 6 provides *statistics*, that is, 
calculations made from the sample of 51,927 people. What we'd like, though, is 
insight into the population *parameters*. You answer the question, "What 
proportion of people in your sample reported being atheists?" with a 
statistic; while the question "What proportion of people on earth would report 
being atheists" is answered with an estimate of the parameter.

The inferential tools for estimating population proportion are analogous to 
those used for means in the last chapter: the confidence interval and the 
hypothesis test.

5.  Write out the conditions for inference to construct a 95% confidence
    interval for the proportion of atheists in the United States in 2012.
    Are you confident all conditions are met?
    
    Each proportion separately must follow a normal model, in order to check conditions for normality, the sample observations must be independent and have at least 10 successes and at least 10 failures.  Additionally, the two samples must be independent of each other.  It seems reasonable to assume these conditions are met since:
    
      - there are 1002 observations in our sample of US residents surveyed in 2012, so if the proportion of aetheists is 5% and non-aetheists are 95% each proportion would meet the success-failure condition
      - the two proportions are independent of each other 
      - the observations can be assumed to be independent since they were randomly selected.

If the conditions for inference are reasonable, we can either calculate
the standard error and construct the interval by hand, or allow the `inference`
function to do it for us.

```{r us-atheism-ci, eval=TRUE, tidy = FALSE}
inference(us12$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")
```

Note that since the goal is to construct an interval estimate for a 
proportion, it's necessary to specify what constitutes a "success", which here 
is a response of `"atheist"`.

Although formal confidence intervals and hypothesis tests don't show up in the 
report, suggestions of inference appear at the bottom of page 7: "In general, 
the error margin for surveys of this kind is $\pm$ 3-5% at 95% confidence".

6.  Based on the R output, what is the margin of error for the estimate of the 
    proportion of atheists in US in 2012?
    
    ```{r}
ME <- 0.0634 - 0.0499
ME
```

    The margin of error is `r round(ME*100, 2)`%.

7.  Using the `inference` function, calculate confidence intervals for the 
    proportion of atheists in 2012 in two other countries of your choice, and 
    report the associated margins of error. Be sure to note whether the 
    conditions for inference are met. It may be helpful to create new data 
    sets for each of the two countries first, and then use these data sets in 
    the `inference` function to construct the confidence intervals.
    
    ##### Germany
    
    It seems reasonable to assume the conditions for inference are met since:
    
      - there are 502 observations in our sample and the proportion of aetheists is 15% and non-aetheists is 85% so each proportion would meet the success-failure condition
      - the two proportions are independent of each other 
      - the observations can be assumed to be independent since they were randomly selected

    
    ```{r}
germany12 <- subset(atheism, nationality == "Germany" & year == "2012")
inference(germany12$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")
ME <- 0.1806 - 0.1494
```
      
    The margin of error is `r round(ME*100, 2)`%.
    
    ##### Italy

    It seems reasonable to assume the conditions for inference are met since:
    
      - there are 987 observations in our sample and the proportion of aetheists is 8% and non-aetheists is 92% so each proportion would meet the success-failure condition
      - the two proportions are independent of each other 
      - the observations can be assumed to be independent since they were randomly selected

    ```{r}
italia12 <- subset(atheism, nationality == "Italy" & year == "2012")
inference(italia12$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")
ME <- 0.097 - 0.08
```

    The margin of error is `r round(ME*100, 2)`%.

## How does the proportion affect the margin of error?

Imagine you've set out to survey 1000 people on two questions: are you female? 
and are you left-handed? Since both of these sample proportions were 
calculated from the same sample size, they should have the same margin of 
error, right? Wrong! While the margin of error does change with sample size, 
it is also affected by the proportion.

Think back to the formula for the standard error: $SE = \sqrt{p(1-p)/n}$. This 
is then used in the formula for the margin of error for a 95% confidence 
interval: $ME = 1.96\times SE = 1.96\times\sqrt{p(1-p)/n}$. Since the 
population proportion $p$ is in this $ME$ formula, it should make sense that 
the margin of error is in some way dependent on the population proportion. We 
can visualize this relationship by creating a plot of $ME$ vs. $p$.

The first step is to make a vector `p` that is a sequence from 0 to 1 with 
each number separated by 0.01. We can then create a vector of the margin of 
error (`me`) associated with each of these values of `p` using the familiar 
approximate formula ($ME = 2 \times SE$). Lastly, we plot the two vectors 
against each other to reveal their relationship.

```{r me-plot, eval=TRUE}
n <- 1000
p <- seq(0, 1, 0.01)
me <- 2 * sqrt(p * (1 - p)/n)
plot(me ~ p, ylab = "Margin of Error", xlab = "Population Proportion")
```

8.  Describe the relationship between `p` and `me`.

    The margin of error `me` increases as the proportion `p` approaches 0.5 and then decreases from 0.5 to 1.

## Success-failure condition

The textbook emphasizes that you must always check conditions before making 
inference. For inference on proportions, the sample proportion can be assumed 
to be nearly normal if it is based upon a random sample of independent 
observations and if both $np \geq 10$ and $n(1 - p) \geq 10$. This rule of 
thumb is easy enough to follow, but it makes one wonder: what's so special 
about the number 10?

The short answer is: nothing. You could argue that we would be fine with 9 or 
that we really should be using 11. What is the "best" value for such a rule of 
thumb is, at least to some degree, arbitrary. However, when $np$ and $n(1-p)$ 
reaches 10 the sampling distribution is sufficiently normal to use confidence 
intervals and hypothesis tests that are based on that approximation.

We can investigate the interplay between $n$ and $p$ and the shape of the 
sampling distribution by using simulations. To start off, we simulate the 
process of drawing 5000 samples of size 1040 from a population with a true 
atheist proportion of 0.1. For each of the 5000 samples we compute $\hat{p}$ 
and then plot a histogram to visualize their distribution.

```{r sim-np, eval=TRUE}
p <- 0.1
n <- 1040
p_hats <- rep(0, 5000)

for(i in 1:5000){
  samp <- sample(c("atheist", "non_atheist"), n, replace = TRUE, prob = c(p, 1-p))
  p_hats[i] <- sum(samp == "atheist")/n
}

hist(p_hats, main = "p = 0.1, n = 1040", xlim = c(0, 0.18))
```

These commands build up the sampling distribution of $\hat{p}$ using the 
familiar `for` loop. You can read the sampling procedure for the first line of 
code inside the `for` loop as, "take a sample of size $n$ with replacement 
from the choices of atheist and non-atheist with probabilities $p$ and $1 - p$,
respectively." The second line in the loop says, "calculate the proportion of 
atheists in this sample and record this value." The loop allows us to repeat 
this process 5,000 times to build a good representation of the sampling 
distribution.

9.  Describe the sampling distribution of sample proportions at $n = 1040$ and 
    $p = 0.1$. Be sure to note the center, spread, and shape.\
    *Hint:* Remember that R has functions such as `mean` to calculate summary
    statistics.
    
    ```{r}
summary(p_hats)
sd(p_hats)
```
    
    The sampling distribution of sample proportions at $n = 1040$ and $p = 0.1$ appears normal in shape with it's center at approximately $p$ or .10, a spread from .07 to .13, and a standard deviation of .009.

10. Repeat the above simulation three more times but with modified sample
    sizes and proportions: for $n = 400$ and $p = 0.1$, $n = 1040$ and 
    $p = 0.02$, and $n = 400$ and $p = 0.02$. Plot all four histograms
    together by running the `par(mfrow = c(2, 2))` command before creating the 
    histograms. You may need to expand the plot window to accommodate the 
    larger two-by-two plot. Describe the three new sampling distributions. 
    Based on these limited plots, how does $n$ appear to affect the 
    distribution of $\hat{p}$? How does $p$ affect the sampling distribution?
    
    ```{r sim-np2, eval=TRUE}
p <- 0.1
n <- 400
p_hats2 <- rep(0, 5000)

for(i in 1:5000){
  samp <- sample(c("atheist", "non_atheist"), n, replace = TRUE, prob = c(p, 1-p))
  p_hats2[i] <- sum(samp == "atheist")/n
}
```

    ```{r sim-np3, eval=TRUE}
p <- 0.02
n <- 1040
p_hats3 <- rep(0, 5000)

for(i in 1:5000){
  samp <- sample(c("atheist", "non_atheist"), n, replace = TRUE, prob = c(p, 1-p))
  p_hats3[i] <- sum(samp == "atheist")/n
}
```

    ```{r sim-np4, eval=TRUE}
p <- 0.02
n <- 400
p_hats4 <- rep(0, 5000)

for(i in 1:5000){
  samp <- sample(c("atheist", "non_atheist"), n, replace = TRUE, prob = c(p, 1-p))
  p_hats4[i] <- sum(samp == "atheist")/n
}
```

    ```{r}
par(mfrow = c(2, 2))
hist(p_hats, main = "p = 0.1, n = 1040", xlim = c(0, 0.18))
hist(p_hats2, main = "p = 0.1, n = 400", xlim = c(0, 0.18))
hist(p_hats3, main = "p = 0.02, n = 1040", xlim = c(0, 0.18))
hist(p_hats4, main = "p = 0.02, n = 400", xlim = c(0, 0.18))
par(mfrow = c(1, 1))
```

    Each distribution has it's center at p.  The two with sample size of 400 have wider spreads while the two with samples of size 1040 have narrower spreads.  The plot with the smaller proportion and larger sample size has the narrowest spread and the distribution with the larger proportion and smaller sample size has the widest spread.  A change in p moves the center of the sampling distribution to the new p. An increase in n narrows the spread while a decrease in n widens the spread of the sampling distribution.

Once you're done, you can reset the layout of the plotting window by using the 
command `par(mfrow = c(1, 1))` command or clicking on "Clear All" above the 
plotting window (if using RStudio). Note that the latter will get rid of all 
your previous plots.

11. If you refer to Table 6, you'll find that Australia has a sample 
    proportion of 0.1 on a sample size of 1040, and that Ecuador has a sample 
    proportion of 0.02 on 400 subjects. Let's suppose for this exercise that 
    these point estimates are actually the truth. Then given the shape of 
    their respective sampling distributions, do you think it is sensible to 
    proceed with inference and report margin of errors, as the reports does?
    
    For Australia this makes sense.  The proportion and sample size correspond to the first of the four plots we did above, so the sample size is sufficiently large enough to have a relatively small margin of error for that proportion.  For Ecuador we don't even meet the success-failure condition.  We would only have 8 successes at .02 p in 400 observations.  So it would not normally make sense to preceed with inference, however, since it is just a small part of a much larger study, it may make sense to report it as long as you highlight the large margin of error, rather than excluding it from your results.
    

* * *
## On your own

The question of atheism was asked by WIN-Gallup International in a similar 
survey that was conducted in 2005. (We assume here that sample sizes have 
remained the same.) Table 4 on page 13 of the report summarizes survey results 
from 2005 and 2012 for 39 countries.

-   Answer the following two questions using the `inference` function. As 
    always, write out the hypotheses for any tests you conduct and outline the
    status of the conditions for inference.

    **a.** Is there convincing evidence that Spain has seen a change in its 
    atheism index between 2005 and 2012?\
    *Hint:* Create a new data set for respondents from Spain. Form
    confidence intervals for the true proportion of athiests in both
    years, and determine whether they overlap.
    
    ##### Spain
    
    ```{r}
spain <- subset(atheism, nationality == "Spain")
    
inference(spain$response, as.factor(spain$year), est = "proportion", type = "ci", method = "theoretical", success = "atheist", order = c("2012","2005"))
```

    Independence, normality and success-failure conditions are sastisfied, so we can use a confidence interval for inference.  However our confidence interval for the difference of proportions for 2005 vs. 2012 includes zero, so we cannot be sure that there is any real difference between 2005 and 2012.  There is not enough convincing evidence to support a claim that Spain has seen a change in its atheism index between 2005 and 2012.

    **b.** Is there convincing evidence that the United States has seen a
    change in its atheism index between 2005 and 2012?
    
    ##### United States
    
    ```{r}
US <- subset(atheism, nationality == "United States")
    
inference(US$response, as.factor(US$year), est = "proportion", type = "ci", method = "theoretical", success = "atheist", order = c("2012","2005"))
```

    Independence, normality and success-failure conditions are sastisfied, so we can use a confidence interval for inference.  Our confidence interval shows a difference of 0.0251 to 0.0547 in the difference between proportions for 2005 vs. 2012, so we can be 95% confident that there is a true difference  between proportions for 2005 and 2012.  There is convincing evidence to support a claim that the US has seen a change in its atheism index between 2005 and 2012 of between 2.5% and 5.5%.

-   If in fact there has been no change in the atheism index in the countries 
    listed in Table 4, in how many of those countries would you expect to 
    detect a change (at a significance level of 0.05) simply by chance?\
    *Hint:* Look in the textbook index under Type 1 error.
    
    Using a .05 significance level, I would expect to detect an error simply by chance in 5% of the cases, and since we have 57 countries in the study, and 57 x .05 = `r 57*.05`, I would expect to get a type 1 error in 2 or 3 countries.

-   Suppose you're hired by the local government to estimate the proportion of 
    residents that attend a religious service on a weekly basis. According to 
    the guidelines, the estimate must have a margin of error no greater than 
    1% with 95% confidence. You have no idea what to expect for $p$. How many 
    people would you have to sample to ensure that you are within the 
    guidelines?\
    *Hint:* Refer to your plot of the relationship between $p$ and margin of 
    error. Do not use the data set to answer this question.
    
    You would need to sample 9604 residents to ensure that your margin of error would never go above 1% if you have no idea what the proportion will be.  
    
    ```{r}
p <- .5
me <- .01
n <- 1.96^2 * (p*(1-p))/me^2
n
```


<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). 
This lab was written for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel.
</div>
